---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# About

`himrr` is an R data package for policy parameters specified by ACA's **H**ealth **I**nsurance **M**arket **R**ating **R**eform.

Premiums in Health Insurance Marketplace are regulated under the Affordable Care Act (ACA).
Specifically, [45 CFR 147.102 - Fair health insurance premiums](https://www.law.cornell.edu/cfr/text/45/147.102) provides a federal default of 

1. Age rating ratios of 3:1,
2. Federally established age curve,
3. Tobacco rating ratios of no more than 1.5:1, 
4. Two tiers: one tier for all adults and a second tier for all children under 21, and
5. Per member rating. 

States can request a standard lower than the federal standard. 
For example, DC, MA, NY, and UT have their own age curve for both individual and small group markets, and NJ has specific age curve for small group market. 

## Source

- [Market Rating Reforms, State Specific Rating Variations, CCIIO; Updated February 26, 2016.](https://www.cms.gov/CCIIO/Programs-and-Initiatives/Health-Insurance-Market-Reforms/state-rating.html)

# Installation

```r
# install.packages("devtools")
devtools::install_github("jjchern/himrr")
```

# Usage

## `himrr`'s Three Tibbles

`himrr` includes three datasets:

- `himrr::default` for federal default, 
- `himrr::im` for state-specific rating variations in individual market, and 
- `himrr::sgm` for state-specific rating variations in small group market.

```{r}
library(dplyr, warn.conflicts = FALSE)
himrr::default
himrr::im
himrr::sgm
```

## State Specific Age Curves for Individual Market

```{r}
himrr::im %>% 
        select(state, age_curve) %>% 
        tidyr::unnest() 
```

## State Specific Family Tiers For Individual Market

```{r}
himrr::im %>% 
        select(state, family_tiers) %>% 
        mutate(state = if_else(family_tiers == "Uniform", "Other States", state)) %>% 
        distinct() %>% 
        knitr::kable()
```

# Reference

- [KFF - Health Insurance & Managed Care](http://kff.org/state-category/health-insurance-managed-care/)
- [The Affordable Care Act: Rating Factor Limitations](http://coventryhealthcare.com/web/groups/public/@cvty_corporate_chc/documents/webcontent/c084481.pdf)
